Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-04-20 reports/manuscript/paper.tex
ENH: Add notes on generalization in discussion section

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}.
As with the original study, it is unclear whether our strong predictive performance based on image properties that would be visible in patient scans or whether it is identifying a new subclinical disease biomarker. 
In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

>>>>

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}.
As with the original study, it is unclear whether our strong predictive performance based on image properties that would be visible in patient scans or whether it is identifying a new subclinical disease biomarker.

In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.
We present age quintile bundle profiles and SGL \protect$\beta$ coefficients for all four datasets in the supplementary material. For age regression, substantial differences between the datasets preclude generalization across all three:
\begin{enumerate*}[%
    label=(\roman*),%
    before=\unskip{: },%
    itemjoin={{, }},%
    itemjoin*={{, and }}]
    \item different acquisition parameters, which can be challenging to harmonize \mbox{\cite{pinto2020harmonization}}
    \item different diffusion models, with DTI for the WH dataset and DKI for the HBN and Cam-CAN datasets
    \item different age ranges and distributions (which is evident in the figure legends for \cref{fig:age-results}), with HBN being a developmental dataset, while WH and Cam-CAN are lifespan maturation datasets
    \item different anatomical extents, with the WH streamlines truncated to remain with the bundle's bounding regions of interest (the default behavior in the legacy \texttt{mAFQ}) and the HBN and Cam-CAN streamlines allowed to retain their full extent from whole-brain tractography (the default behavior in \texttt{pyAFQ}).
\end{enumerate*}
In a biological characteristic that has widespread effects in the brain, validity is difficult to assess \mbox{\cite{lermausabiaga2019replication}, so it is both unsurprising and vexing that we can do so well at predicting within each dataset, and yet do poorly in interpreting informative features across instruments.}

### 2021-04-20 reports/manuscript/paper_changes.tex
ENH: Add notes on generalization in discussion section

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}.
\added{As with the original study, it is unclear whether our strong predictive performance based on image properties that would be visible in patient scans or whether it is identifying a new subclinical disease biomarker.} 
In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

>>>>

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}.
\added{As with the original study, it is unclear whether our strong predictive performance based on image properties that would be visible in patient scans or whether it is identifying a new subclinical disease biomarker.}

In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.
\added{We present age quintile bundle profiles and SGL \protect$\beta$ coefficients for all four datasets in the supplementary material. For age regression, substantial differences between the datasets preclude generalization across all three:}
\begin{enumerate*}[%
    label=(\roman*),%
    before=\unskip{: },%
    itemjoin={{, }},%
    itemjoin*={{, and }}]
    \item \added{different acquisition parameters, which can be challenging to harmonize \mbox{\cite{pinto2020harmonization}}}
    \item \added{different diffusion models, with DTI for the WH dataset and DKI for the HBN and Cam-CAN datasets}
    \item \added{different age ranges and distributions (which is evident in the figure legends for \cref{fig:age-results}), with HBN being a developmental dataset, while WH and Cam-CAN are lifespan maturation datasets}
    \item \added{different anatomical extents, with the WH streamlines truncated to remain with the bundle's bounding regions of interest (the default behavior in the legacy \texttt{mAFQ}) and the HBN and Cam-CAN streamlines allowed to retain their full extent from whole-brain tractography (the default behavior in \texttt{pyAFQ}).}
\end{enumerate*}
\added{In a biological characteristic that has widespread effects in the brain, validity is difficult to assess \mbox{\cite{lermausabiaga2019replication}}, so it is both unsurprising and vexing that we can do so well at predicting within each dataset, and yet do poorly in interpreting informative features across instruments.}

### 2021-04-20 reports/manuscript/supplemental.tex
ENH: Add notes on generalization in discussion section

\Cref{fig:hbn-bp:fa,fig:hbn-bp:md} show the bundle profiles and regression
coefficients for the HBN dataset. Like the WH dataset, the $\hat{\beta}$
coefficients are distributed widely through the brain and SGL behaves more
like the lasso than the group lasso. In contrast the the WH results,
the bundle profiles show different behaviors. For example the SLFL and SLFR
bundle profiles in \cref{fig:wh-bp:md} and \cref{fig:hbn-bp:md} have
different concavity. This is unsurprising, however, given the differences
between these datasets
\begin{enumerate*}[%
    label=(\roman*),%
    before=\unskip{: },%
    itemjoin={{, }},%
    itemjoin*={{, and }}]
    \item different diffusion models, with DTI for the WH dataset and DKI for
    the HBN dataset
    \item different age ranges and distributions (which is evident in the
    figure legends), with HBN being a developmental dataset, while WH is a
    lifespan maturation dataset
    \item different anatomical extents, with the WH streamlines truncated to
    remain with the bundle's bounding regions of interest (the default
    behavior in the legacy \texttt{mAFQ}) and the HBN streamlines allowed
    to retain their full extent from whole-brain tractography (the default
    behavior in \texttt{pyAFQ}).
\end{enumerate*}
Thus, one should use caution when comparing bundle profiles and $\hat{\beta}$
coefficients between the WH, HBN, and Cam-CAN models.
The parsimony pitfall is most evident in the UNCL, UNCR, ARCL, SLFL, and
SLFR bundles in \cref{fig:hbn-bp:md}.

>>>>

\Cref{fig:hbn-bp:fa,fig:hbn-bp:md} show the bundle profiles and regression
coefficients for the HBN dataset. Like the WH dataset, the $\hat{\beta}$
coefficients are distributed widely through the brain and SGL behaves more
like the lasso than the group lasso. In contrast the the WH results,
the bundle profiles show different behaviors. For example the SLFL and SLFR
bundle profiles in \cref{fig:wh-bp:md} and \cref{fig:hbn-bp:md} have
different concavity. This is unsurprising, however, given the differences
between these datasets
\begin{enumerate*}[%
    label=(\roman*),%
    before=\unskip{: },%
    itemjoin={{, }},%
    itemjoin*={{, and }}]
    \item different diffusion models, with DTI for the WH dataset and DKI for
    the HBN and Cam-CAN datasets
    \item different age ranges and distributions (which is evident in the
    figure legends), with HBN being a developmental dataset, while WH and Cam-CAN are
    lifespan maturation datasets
    \item different anatomical extents, with the WH streamlines truncated to
    remain with the bundle's bounding regions of interest (the default
    behavior in the legacy \texttt{mAFQ}) and the HBN and Cam-CAN streamlines
    allowed to retain their full extent from whole-brain tractography (the
    default behavior in \texttt{pyAFQ}).
\end{enumerate*}
Thus, one should use caution when comparing bundle profiles and $\hat{\beta}$
coefficients between the WH, HBN, and Cam-CAN models.
The parsimony pitfall is most evident in the UNCL, UNCR, ARCL, SLFL, and
SLFR bundles in \cref{fig:hbn-bp:md}.
